\bf Calendar-based graphics for visualizing people's daily schedules

### 2018-10-15 main.Rmd
switched flinders to birrarung marr

Figure \ref{fig:flinders-2016} shows the line glyphs framed in the monthly calendar over the year 2016. This is achieved by the `frame_calendar` function, which computes the coordinates on the calendar for the input data variables. These can then be plotted using the usual **ggplot2** package [@R-ggplot2] functions. All of the grammar of graphics [@wilkinson2006grammar; @wickham2009ggplot2] can be applied.

```{r flinders-2016, fig.height = 7, fig.cap = "The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs. The arrangement of the data into a $3 \\times 4$ monthly grid represents all the traffic in 2016. The disparities between week day and weekend along with public holiday are immediately apparent."}
```

@@

Regarding the monthly calendar, the major reference lines separate every month panel and the minor ones separate every cell, represented by the thick and thin lines in Figure \ref{fig:flinders-2016}, respectively. The major reference lines are placed surrounding every month block: for each $m$, the vertical lines are determined by $\min{(x)}$ and $\max{(x)}$; for each $n$, the horizontal lines are given by $\min{(y)}$ and $\max{(y)}$. The minor reference lines are only placed on the left side of every cell: for each $i$, the vertical division is $\min{(x)}$; for each $j$, the horizontal is $\min{(y)}$.

@@

It is assumed that the `data` is in tidy format [@wickham2014tidy], and `x`, `y` are the variables that will be mapped to the horizontal and vertical axes in each cell. For example, the `x` is the time of the day, and `y` is the count (Figure \ref{fig:flinders-2016}). The `date` argument specifies the date variable used to construct the calendar layout.

@@

When `polar = TRUE`, a polar transformation is carried out on the data. The computation is similar to the one described in @Wickham2012glyph. Figure \ref{fig:flinders-polar} shows star plots embedded in the monthly calendar layout, which is equivalent to Figure \ref{fig:flinders-2016} placed in polar coordinates.

```{r flinders-polar, fig.height = 7, fig.cap = "Figure \\ref{fig:flinders-2016} in circular layout, which is referred to as star plots. The daily periodicity on work days are clearly visible."}
```

@@

With local scaling, the overall variation gives way to the individual shape. Figure \ref{fig:flinders-free} shows the same data as Figure \ref{fig:flinders-2016} scaled locally using `scale = "free"`. The daily trends are magnified.

@@

```{r flinders-free, fig.height = 7, fig.cap = "Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days. The individual shape on a single day becomes more distinctive, however it is impossible to compare the size of peaks between days."}
```

@@

Plots can be layered. The comparison of sensors can be done by overlaying plot the values for each (Figure \ref{fig:overlay}). Differences between the pedestrian patterns for Flinders Street and Flagstaff train stations can be seen. Both Flinders Street and Flagstaff train stations exhibit strong  commuters patterns, however, Flinders Street has higher pedestrian counts during the weekends and public holidays. This suggests that Flagstaff Station has limited functionality on non-work days, but other activities occur around the Flinders Street Station. From Figure \ref{fig:overlay} it can be seen that the State Library has a similar temporal trend to Flinders Street Station on non-work days. The nighttime events, such as White Night and New Year's Eve, have barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to Flinders Street Station and the State Library.

@@

The `frame_calendar` function is not constrained to line plots. The full range of plotting capabilities in **ggplot2** is essentially available. Figure \ref{fig:scatterplot} shows a lag scatterplot at Flinders Street Station, where the lagged hourly count is assigned to the `x` argument and the current hourly count to the `y` argument. This figure is organized in the daily calendar layout. Figure \ref{fig:scatterplot} indicates two primary patterns, strong autocorrelation on weekends, and weaker autocorrelation on work days. At the higher counts, on week days, the next hour sees possibly substantial increase or decrease in counts, essentially revealing a bimodal distribution of consecutive counts, as supported by Figure \ref{fig:flinders-2016}.

```{r scatterplot, fig.height = 5, fig.cap = "Lag scatterplot in the daily calendar layout. Each hour's count is plotted against previous hour's count at Flinders Street Station to demonstrate the autocorrelation at lag 1. The correlation between them is more consistent on non-work days than work days."}
```

>>>>

Figure \ref{fig:fs-2016} shows the line glyphs framed in the monthly calendar over the year 2016. This is achieved by the `frame_calendar` function, which computes the coordinates on the calendar for the input data variables. These can then be plotted using the usual **ggplot2** package [@R-ggplot2] functions. All of the grammar of graphics [@wilkinson2006grammar; @wickham2009ggplot2] can be applied.

```{r fs-2016, fig.height = 7, fig.cap = "The calendar-based display of hourly foot traffic at Flagstaff Station using line glyphs. The arrangement of the data into a $3 \\times 4$ monthly grid represents all the traffic in 2016. The disparities between week day and weekend along with public holiday are immediately apparent."}
```

@@

Regarding the monthly calendar, the major reference lines separate every month panel and the minor ones separate every cell, represented by the thick and thin lines in Figure \ref{fig:fs-2016}, respectively. The major reference lines are placed surrounding every month block: for each $m$, the vertical lines are determined by $\min{(x)}$ and $\max{(x)}$; for each $n$, the horizontal lines are given by $\min{(y)}$ and $\max{(y)}$. The minor reference lines are only placed on the left side of every cell: for each $i$, the vertical division is $\min{(x)}$; for each $j$, the horizontal is $\min{(y)}$.

@@

It is assumed that the `data` is in tidy format [@wickham2014tidy], and `x`, `y` are the variables that will be mapped to the horizontal and vertical axes in each cell. For example, the `x` is the time of the day, and `y` is the count (Figure \ref{fig:fs-2016}). The `date` argument specifies the date variable used to construct the calendar layout.

@@

When `polar = TRUE`, a polar transformation is carried out on the data. The computation is similar to the one described in @Wickham2012glyph. Figure \ref{fig:fs-polar} shows star plots embedded in the monthly calendar layout, which is equivalent to Figure \ref{fig:fs-2016} placed in polar coordinates.

```{r fs-polar, fig.height = 7, fig.cap = "Figure \\ref{fig:fs-2016} in circular layout, which is referred to as star plots. The daily periodicity on work days are clearly visible."}
```

@@

With local scaling, the overall variation gives way to the individual shape. Figure \ref{fig:fs-free} shows the same data as Figure \ref{fig:fs-2016} scaled locally using `scale = "free"`. The daily trends are magnified.

@@

```{r fs-free, fig.height = 7, fig.cap = "Line glyphs on the calendar format showing hourly foot traffic at Flagstaff Station, scaled over all the days. The individual shape on a single day becomes more distinctive, however it is impossible to compare the size of peaks between days."}
```

@@

Plots can be layered. The comparison of sensors can be done by overlaying plot the values for each (Figure \ref{fig:overlay}). Differences between the pedestrian patterns at these sensors can be seen. Flagstaff Station exhibits strong commuters patterns, with fewer pedestrian counts during the weekends and public holidays. This suggests that Flagstaff Station has limited functionality on non-work days. From Figure \ref{fig:overlay} it can be seen that the State Library has a similar temporal trend to Flagstaff Station on non-work days. The nighttime events, such as White Night and New Year's Eve, have barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to Flagstaff Station and the State Library.

@@

The `frame_calendar` function is not constrained to line plots. The full range of plotting capabilities in **ggplot2** is essentially available. Figure \ref{fig:scatterplot} shows a lag scatterplot at Flagstaff Station, where the lagged hourly count is assigned to the `x` argument and the current hourly count to the `y` argument. This figure is organized in the daily calendar layout. Figure \ref{fig:scatterplot} indicates two primary patterns, strong autocorrelation on weekends, and weaker autocorrelation on work days. At the higher counts, on week days, the next hour sees possibly substantial increase or decrease in counts, essentially revealing a bimodal distribution of consecutive counts, as supported by Figure \ref{fig:fs-2016}.

```{r scatterplot, fig.height = 5, fig.cap = "Lag scatterplot in the daily calendar layout. Each hour's count is plotted against previous hour's count at Flagstaff Station to demonstrate the autocorrelation at lag 1. The correlation between them is more consistent on non-work days than work days."}
```

### 2018-10-15 main.tex
switched flinders to birrarung marr

Figure \ref{fig:flinders-2016} shows the line glyphs framed in the
monthly calendar over the year 2016. This is achieved by the
\texttt{frame\_calendar} function, which computes the coordinates on the
calendar for the input data variables. These can then be plotted using
the usual \textbf{ggplot2} package \citep{R-ggplot2} functions. All of
the grammar of graphics \citep{wilkinson2006grammar, wickham2009ggplot2}
can be applied.

@@

{\centering \includegraphics[width=\textwidth]{figure/flinders-2016-1}

@@

\caption{The calendar-based display of hourly foot traffic at Flinders Street Station using line glyphs. The arrangement of the data into a $3 \times 4$ monthly grid represents all the traffic in 2016. The disparities between week day and weekend along with public holiday are immediately apparent.}\label{fig:flinders-2016}
\end{figure}

@@

Regarding the monthly calendar, the major reference lines separate every
month panel and the minor ones separate every cell, represented by the
thick and thin lines in Figure \ref{fig:flinders-2016}, respectively.
The major reference lines are placed surrounding every month block: for
each \(m\), the vertical lines are determined by \(\min{(x)}\) and
\(\max{(x)}\); for each \(n\), the horizontal lines are given by
\(\min{(y)}\) and \(\max{(y)}\). The minor reference lines are only
placed on the left side of every cell: for each \(i\), the vertical
division is \(\min{(x)}\); for each \(j\), the horizontal is
\(\min{(y)}\).

@@

It is assumed that the \texttt{data} is in tidy format
\citep{wickham2014tidy}, and \texttt{x}, \texttt{y} are the variables
that will be mapped to the horizontal and vertical axes in each cell.
For example, the \texttt{x} is the time of the day, and \texttt{y} is
the count (Figure \ref{fig:flinders-2016}). The \texttt{date} argument
specifies the date variable used to construct the calendar layout.

@@

When \texttt{polar\ =\ TRUE}, a polar transformation is carried out on
the data. The computation is similar to the one described in
\citet{Wickham2012glyph}. Figure \ref{fig:flinders-polar} shows star
plots embedded in the monthly calendar layout, which is equivalent to
Figure \ref{fig:flinders-2016} placed in polar coordinates.

@@

{\centering \includegraphics[width=\textwidth]{figure/flinders-polar-1}

@@

\caption{Figure \ref{fig:flinders-2016} in circular layout, which is referred to as star plots. The daily periodicity on work days are clearly visible.}\label{fig:flinders-polar}
\end{figure}

@@

With local scaling, the overall variation gives way to the individual
shape. Figure \ref{fig:flinders-free} shows the same data as Figure
\ref{fig:flinders-2016} scaled locally using \texttt{scale\ =\ "free"}.
The daily trends are magnified.

@@

{\centering \includegraphics[width=\textwidth]{figure/flinders-free-1}

@@

\caption{Line glyphs on the calendar format showing hourly foot traffic at Flinders Street Station, scaled over all the days. The individual shape on a single day becomes more distinctive, however it is impossible to compare the size of peaks between days.}\label{fig:flinders-free}
\end{figure}

@@

The \texttt{frame\_calendar} function is not constrained to line plots.
The full range of plotting capabilities in \textbf{ggplot2} is
essentially available. Figure \ref{fig:scatterplot} shows a lag
scatterplot at Flinders Street Station, where the lagged hourly count is
assigned to the \texttt{x} argument and the current hourly count to the
\texttt{y} argument. This figure is organized in the daily calendar
layout. Figure \ref{fig:scatterplot} indicates two primary patterns,
strong autocorrelation on weekends, and weaker autocorrelation on work
days. At the higher counts, on week days, the next hour sees possibly
substantial increase or decrease in counts, essentially revealing a
bimodal distribution of consecutive counts, as supported by Figure
\ref{fig:flinders-2016}.

@@

\caption{Lag scatterplot in the daily calendar layout. Each hour's count is plotted against previous hour's count at Flinders Street Station to demonstrate the autocorrelation at lag 1. The correlation between them is more consistent on non-work days than work days.}\label{fig:scatterplot}
\end{figure}

>>>>

Figure \ref{fig:fs-2016} shows the line glyphs framed in the monthly
calendar over the year 2016. This is achieved by the
\texttt{frame\_calendar} function, which computes the coordinates on the
calendar for the input data variables. These can then be plotted using
the usual \textbf{ggplot2} package \citep{R-ggplot2} functions. All of
the grammar of graphics \citep{wilkinson2006grammar, wickham2009ggplot2}
can be applied.

@@

{\centering \includegraphics[width=\textwidth]{figure/fs-2016-1}

@@

\caption{The calendar-based display of hourly foot traffic at Flagstaff Station using line glyphs. The arrangement of the data into a $3 \times 4$ monthly grid represents all the traffic in 2016. The disparities between week day and weekend along with public holiday are immediately apparent.}\label{fig:fs-2016}
\end{figure}

@@

Regarding the monthly calendar, the major reference lines separate every
month panel and the minor ones separate every cell, represented by the
thick and thin lines in Figure \ref{fig:fs-2016}, respectively. The
major reference lines are placed surrounding every month block: for each
\(m\), the vertical lines are determined by \(\min{(x)}\) and
\(\max{(x)}\); for each \(n\), the horizontal lines are given by
\(\min{(y)}\) and \(\max{(y)}\). The minor reference lines are only
placed on the left side of every cell: for each \(i\), the vertical
division is \(\min{(x)}\); for each \(j\), the horizontal is
\(\min{(y)}\).

@@

It is assumed that the \texttt{data} is in tidy format
\citep{wickham2014tidy}, and \texttt{x}, \texttt{y} are the variables
that will be mapped to the horizontal and vertical axes in each cell.
For example, the \texttt{x} is the time of the day, and \texttt{y} is
the count (Figure \ref{fig:fs-2016}). The \texttt{date} argument
specifies the date variable used to construct the calendar layout.

@@

When \texttt{polar\ =\ TRUE}, a polar transformation is carried out on
the data. The computation is similar to the one described in
\citet{Wickham2012glyph}. Figure \ref{fig:fs-polar} shows star plots
embedded in the monthly calendar layout, which is equivalent to Figure
\ref{fig:fs-2016} placed in polar coordinates.

@@

{\centering \includegraphics[width=\textwidth]{figure/fs-polar-1}

@@

\caption{Figure \ref{fig:fs-2016} in circular layout, which is referred to as star plots. The daily periodicity on work days are clearly visible.}\label{fig:fs-polar}
\end{figure}

@@

With local scaling, the overall variation gives way to the individual
shape. Figure \ref{fig:fs-free} shows the same data as Figure
\ref{fig:fs-2016} scaled locally using \texttt{scale\ =\ "free"}. The
daily trends are magnified.

@@

{\centering \includegraphics[width=\textwidth]{figure/fs-free-1}

@@

\caption{Line glyphs on the calendar format showing hourly foot traffic at Flagstaff Station, scaled over all the days. The individual shape on a single day becomes more distinctive, however it is impossible to compare the size of peaks between days.}\label{fig:fs-free}
\end{figure}

@@

The \texttt{frame\_calendar} function is not constrained to line plots.
The full range of plotting capabilities in \textbf{ggplot2} is
essentially available. Figure \ref{fig:scatterplot} shows a lag
scatterplot at Flagstaff Station, where the lagged hourly count is
assigned to the \texttt{x} argument and the current hourly count to the
\texttt{y} argument. This figure is organized in the daily calendar
layout. Figure \ref{fig:scatterplot} indicates two primary patterns,
strong autocorrelation on weekends, and weaker autocorrelation on work
days. At the higher counts, on week days, the next hour sees possibly
substantial increase or decrease in counts, essentially revealing a
bimodal distribution of consecutive counts, as supported by Figure
\ref{fig:fs-2016}.

@@

\caption{Lag scatterplot in the daily calendar layout. Each hour's count is plotted against previous hour's count at Flagstaff Station to demonstrate the autocorrelation at lag 1. The correlation between them is more consistent on non-work days than work days.}\label{fig:scatterplot}
\end{figure}
